\chapter{Introduction\label{cha:chapter1}}
There is a vast variety of different object detection algorithms reaching from feature-based to template-based methods, and more recently, also deep-learning-based approaches. Especially the latter have conquered the field very rapidly and have consistently improved the state of the art in many established object detection benchmarks. However, this improvement has not yet arrived at the shop floors of the manufacturing companies. This is in part due to characteristics of the method, such as the large amount of required training data. However, another reason is that the current machine vision infrastructure is not flexible enough to keep track of the dynamic development in the field of object detection. Deploying or even only trying out a new method typically requires a certain investment in cost and time that has to be justified very well. This workflow can be greatly accelerated by using \textit{Service-Oriented Architectures} (SOA).

The goal of this thesis is to simplify the process of programming an object detector and integrating it into the manufacturing process. In fact, the great aim is to not program anything at all – instead, a framework shall be proposed that allows generating an object detection service from a single example image.
The service should detect the object with a specified method and should automatically have the appropriate interfaces for convenient integration into SOA. The key advantage is that new methods can be tested and deployed without any effort, as we can generate and deploy a service which has the same interface as our old one but an updated method.


 
 TODO: Die Einleitung muss ein wenig ausführlicher sein. Gönn dir ruhig jeweils einen Abschnitt zu: Kontext, konkrete Problemstellung (inkl. der wissenschaftlichen Herausforderungen), und Herangehensweise in deiner Arbeit.
Du kannst als Motivation auch bereits kurz darstellen wie Bildverarbeitung derzeit in der Produktion integriert ist. Peile so 3 Seiten an, selbst dein Abstract ist gerade ausführlicher als die Einleitung.

Und: Kritische Aussagen müssen mit Quellen belegt werden.